Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-06-05 Manuscript/manuscript.Rmd
results edits

<!--We then went on the evaluate the effect of class imbalance on model performance.--> In order to set a baseline in performance, we first evaluated the least granular random forest classifier, that is one implemented using OC predictor types, without data re-sampling. We found that models built on data with fully imbalanced classes exhibited poor MCCs across all resolutions (Figure 4A). We gradually increased the complexity in data level characteristics by assessing models built on OP predictor types, while still not performing any data resampling. Here, the results were similar to that of OC types (Figure 4B). However, when including distance-type predictors, models were found to exhibit a marked improvement in MCC compared to the other two types (Figure 4C). Results were found to be consistently poor across resolutions and predictor types, when implemented on the K562 cell line (Supplementary Figure 3). Furthermore, we found that conventional metrics of performance (i.e. accuracy and AUC) were less sensitive to models built on imbalanced data, regardless of cell line, resolution, or predictor-type (Supplementary Tables 6-7).

@@

When assessing performances for models built on balanced data, we found that SMOTE re-sampling provided the optimal MCC consistently across both predictor type and resolution (Figure 4). Specifically, the most improvement in model performance [?? Over what ??] was seen when combining SMOTE re-sampling with distance-type predictors. This was also the optimal combination for the K562 cell line (Supplementary Figure 3). These results imply that a best performing model, one that is able to best discriminate between genomic bins that are near versus far away from TAD boundaries [?? Is this what you are doing? Or are you identifying TAD boundaries themselves ??], is one that is built on 10 kb resolution data, with distance-type predictors, and incorporates both random under- and over-sampling. Moreover, there was a clear degredation in model performance as data resolution decreased. Likewise, performances exhibited much more variability across chromosomes when built on lower resolutions.

>>>>

In order to set a baseline in performance, we first evaluated the least granular random forest classifier, that is one implemented using OC predictor types, without data re-sampling. We found that models built on data with fully imbalanced classes exhibited poor MCCs across all resolutions (Figure 4A). We gradually increased the complexity in data level characteristics by assessing models built on OP predictor types, while still not performing any data resampling. Here, the results were similar to that of OC types (Figure 4B). However, when including distance-type predictors, models were found to exhibit a marked improvement in MCC compared to the other two types (Figure 4C). Results were found to be consistently poor across resolutions and predictor types, when implemented on the K562 cell line (Supplementary Figure 3). Furthermore, we found that conventional metrics of performance (i.e. accuracy and AUC) were less sensitive to models built on imbalanced data, regardless of cell line, resolution, or predictor-type (Supplementary Tables 6-7).

@@

When assessing performances for random forest classifiers built on balanced data, we found that SMOTE re-sampling provided the optimal MCC consistently outperforming models built on ROS or RUS, across both predictor type and resolution (Figure 4). Specifically, the most improvement in model performance was seen when combining SMOTE re-sampling with distance-type predictors. This was also the optimal combination for the K562 cell line (Supplementary Figure 3). Interestingly, we found that there was a degredation in model performance as data resolution decreased. Chromosome specific models built on 10 kb data resolutions presented with the highest MCC values. Additionally, performances tended to be much more variable between chromosomes at lower resolutions. Moreover, this decreasing trend in performance was much more evident under the optimal combination of data level characteristics. These results imply that an optimally performing random forest classifier, one that is able to best discriminate between genomic bins that are near versus far away from TAD boundaries, is one that is built on 10 kb resolution data, with distance-type predictors, and incorporates both random under- and over-sampling.
